---
title: "Assignment 3 - CRD"
author: "Jyoti Neupane"  
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: false
    theme: cerulean
---

# Introduction  
This assignment will ask you to conduct a full CRD analysis, from data import through publication ready plot.  

The different sections representing what we did in class were left in the script to provide you a scaffold to start from.  

# Instructions  
  - Move this script into the `code` subfolder of your `03_crd` project folder  
  - Move the file `wheat_nk_bamyan.csv` into the `data` folder of your `03_crd` project folder  
  - On this script, fill your name under the `title` section of the YAML  

# a) Setup  
Here is where we load the packages we will use.  
Make sure to load each package in the chunk below.  
```{r setup, message = FALSE, warning =FALSE}

library(tidyverse) # for data wrangling and plotting
library(car) # for Anova function
library(broom) # for model residuals extraction
library(emmeans) # for model mean extraction
library(multcomp) # for pairwise comparison letter display
library(readr)
library(dplyr) # wrangling
library(tidyr)
library(janitor)
library("multcompView")
```

```

Now, import the data set `wheat_nk_bamyan.csv` in the chunk below. 
```
```{r data import}
#getwd()
crd_assign = read_csv("../Data/wheat_nk_bamyan.csv")
crd_assign

```

The **objective of this study** was to assess the effect of different N and K rates on wheat yield.  

For that, a study was conducted using a completely randomized design as the experimental design with four replicates. The treatment design was a 3 N x 3 K rate factorial, for a total of 9 treatments.  


# b) EDA tables  
Use this space to conduct exploratory data analysis of the imported data, in the form of tables.  

```{r summary} 
summary(crd_assign)
glimpse(crd_assign)

```

# c) Wrangling  
Here, make sure to wrangle your data set with all needed steps. Refer back to class code if you need a refresher on what was needed.  

```{r crd_asw}
crd_asw = crd_assign %>%
  mutate(rep= factor(rep),
          nrate_kgha = factor(nrate_kgha),
         krate_kgha = factor(krate_kgha)
          
          )%>%
  mutate(trtname = paste0(nrate_kgha,"+", krate_kgha))

crd_asw
summary(crd_asw)
```
# d) EDA plots  
Now, explore the wrangled data using plots.  
Make sure to explore plots showing the main effects and the interaction.  
```{r n_rate boxplot}
ggplot(crd_asw, aes(x=nrate_kgha,
                    y= yield_kgha,
                    color= nrate_kgha)) +
  geom_boxplot()+
  geom_jitter()+
  theme(legend.position="none")


```

##Interpretation for n_rate boxplot: From the plot above we can say the wheat yield is increasing with increase nitrogen rate.

```{r k_rate boxplot}
ggplot(crd_asw, aes(x = krate_kgha, 
                    y = yield_kgha,
                    color = krate_kgha)) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none")
```

##Interpretation for n_rate boxplot: Similar to the n_rate boxplot, wheat yield increase with increasing the rate of potassium.

```{r n_k interation boxplot}
ggplot(crd_asw, aes(x=nrate_kgha,
                    y= yield_kgha,
                    color=nrate_kgha
                    
                    ))+
  geom_boxplot()+
  geom_jitter()+
  facet_grid(.~krate_kgha)


```

##Interpretation for n_k interaction boxplot: From the graph above we can see that wheat yield is increasing with nitrogen level in every levels of potassium, however, the overall yield is maximum at 60 kg_ha K rate and is increasing with increase k rate. Thus, it indicate that there might be insignificant interaction between N and K. However, to make any conclusion we need to do ANOVA and refer to P-value.  

## Question 1) Below each plot of this section, write a sentence explaining your visual interpretation of the plot.  

# e) Statistical model  
In this section, run the appropriate model.  
Remember to set the contrast type to sum-to-zero (the default in R is set-to-zero, which is not what we want).  

Also, remember to specify all model terms needed for the treatment design.  
```{r model}
# Changing to sum-to-zero contrast
options(contrasts = c("contr.sum", "contr.poly"))

# Model fitting
crd_m1 = lm(yield_kgha~nrate_kgha+krate_kgha+nrate_kgha:krate_kgha,
             data=crd_asw)

# Summary
summary(crd_m1)
```

# f) ANOVA table  
Use this section to check the ANOVA table, ensuring you use type 3 sum-of-squares.
```{r}
Anova(crd_m1, type=3)
```

## Question 2) Based on the ANOVA table above, for which effect(s) in the ANOVA should you extract means and perform pair-wise comparisons? Explain why.  
##Question 2 Answer: Based on ANOVA table above n rate and k rate has significant effects on yield, however, the their interaction is insignificant thus we should extract means for main effect (i.e. n rate and k rate) separately ans don't need to do interaction means

# g) Linear model assumptions  
Use this section to first compute your residuals, including studentdized residuals.  

```{r}
crd_resid = augment(crd_m1) %>%
  mutate(.studresid= rstudent(crd_m1))

crd_resid
```
## Question 3) What are the linear model assumptions that we need to check for? Create subsections (using ##) below for each of these assumptions, and write code to produce plots to check for these assumptions. For each assumption, write down your interpretation of the produced plots, and your thoughts if the assumptions hold or not.  

##The linear model assumptions are given below: 
- Residual independence  
- Residual variance homogeneity  
- Residual normality 
- Outlier detection

## Residual independence: The residual plot use as a diagnostic tool to check for violations of the residual independence assumption. 

```{r }
ggplot(crd_resid, aes(x= .fitted,
                      y= .studresid))+
  geom_point(shape=21,
             fill="black",
             size=4,
             alpha=.7
             )+
  
  geom_hline(yintercept = c(-3,0,3), color="orange")+
  geom_smooth()
```

##Interpretation: The residual independence assumption is not violated based on my observation.

##Residual homoscedasticity 

```{r}
ggplot(crd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="orange")+
  geom_point(shape = 21,
             fill = "black", 
             size = 4,
             alpha = .7)+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "orange")+
  theme_bw()
```

##Interpretation: In my opinion the contant variance assumption is not violted as I dont see any pattern or shape.


##Residual normality  

```{r}
ggplot(crd_resid, aes(sample=.studresid))+
  stat_qq()+
  stat_qq_line()
```

##Interpretation: Data looks normal to me as the most of the data point fall on the line

##Density plot
```{r}
ggplot(crd_resid, aes(x= .studresid))+
  geom_density()+
  scale_x_continuous(breaks=c(-3,0,3), limits=c(-3,3))
```

##Interpretation: The density plot looks very similar to standard curve and mean is close to zero.


##Residual outliers  

```{r}
ggplot(crd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="orange")+
  geom_point(shape = 21,
             fill = "black", 
             size = 4,
             alpha = .7)+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "orange")+
  theme_bw()
```

##Interpretation: All residuals fall within [-3,3] on a studentized residual scale.So there is no outlier in this data set.
# As all the required model assumptions have been checked and met, we can proceed to using the model for inference.


# h) Model means  
Use this section to extract model means according to your decision of the ANOVA table effects.  
```{r N means}
crd_N_means= emmeans(crd_m1,
                       ~nrate_kgha
                       )
crd_N_means


```
```{r K means}
crd_K_means= emmeans(crd_m1,
                       ~krate_kgha
                       )
crd_K_means

```
# i) Pairwise comparisons  
Use this section to perform pair-wise comparison of the means extracted above.  
```{r N means comparison}
crd_cld_N <- cld(crd_N_means,
                   reversed = T, 
                   Letters= letters,
                   adjust= "none"
                   )

crd_cld_N
```

Make sure to reverse the order from high to low, use letters instead of numbers, and no adjustment.  
```{r K means comparision}
crd_cld_K <- cld(crd_K_means,
                   reversed = T, 
                   Letters= letters,
                   adjust= "none"
                   )

crd_cld_K
```

```{r}
crd_cld_selected_N <- crd_cld_N %>%
  as.data.frame() %>%
  mutate(letter= trimws(.group))

crd_cld_selected_N
```
```{r}
crd_cld_selected_K <- crd_cld_K %>%
  as.data.frame() %>%
  mutate(letter= trimws(.group))

crd_cld_selected_K

```

# g) Final plot(s)  
Produce final plot(s) to show your results.  
Make sure to include both **raw data** and **statistical model summary (i.e., letter separation)**.  

```{r N plot}
ggplot(mapping = aes(fill = nrate_kgha))+
  # Raw data and boxplots  
geom_boxplot(data= crd_asw,
             aes(x= nrate_kgha, y=yield_kgha),
             alpha = .8) +
  
  # Adding letters

  geom_label(data= crd_cld_selected_N,
             aes(x= nrate_kgha,
                 y= emmean, 
                 label= letter,
                 fill = "yellow"
                 )) +
   geom_jitter(data = crd_asw,
              aes(x= nrate_kgha,
                  y= yield_kgha),
              shape = 21,
              alpha = .6,
              size = 4) +
   scale_fill_viridis_d() +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
        
        )

```

##Interpretation for nitrogen plot: The different color of box indicates differetn levels of nitrogen treatment and the yield is increasing with increasing nitrogen level. More specifically, there is no statstically significant difference in yield at 100 n rate and 200 n rate whereas there is significant difference at 0 n rate compared to other two rates.  

```{r K plot}
ggplot(mapping = aes(fill = krate_kgha))+
  # Raw data and boxplots  
geom_boxplot(data= crd_asw,
             aes(x= krate_kgha, y=yield_kgha),
             alpha = .8) +
  
  # Adding letters

  geom_label(data= crd_cld_selected_K,
             aes(x= krate_kgha,
                 y= emmean, 
                 label= letter,
                 fill = "yellow"
                 )) +
   geom_jitter(data = crd_asw,
              aes(x= krate_kgha,
                  y= yield_kgha),
              shape = 21,
              alpha = .6,
              size = 4) +
   scale_fill_viridis_d() +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
        
        )



```

##Interpretation for nitrogen plot: The different color of box indicates differetn levels of potassium treatment and the yield is increasing with increasing potassium level. More specifically, there is no statstically significant difference in yield at 30 k rate and 60 k rate whereas there is significant difference at 0 k rate compared to other two rates. 


Be mindful of your plot to ensure you are creating it in an intuitive way for your audience in relation to interpreting the letter separation structure you used.  

In class, we did not have time to make it publication-ready (e.g., a plot looking professional and ready to be included in a manuscript).

Here, make sure you deliver final plot(s) that is/are publication ready.  

Make sure to provide an interpretation for the plot(s) created, including which treatment levels were significantly higher/lower.


# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `Assignment-03-crd-LASTNAME.html`.    

For ex., mine would be `Assignment-03-crd-Bastos.html`.

Send the **html** file to **lmbastos@uga.edu** by Feb 15th 11:59 pm.  
  




